---
title: "Wrgl_FishCounts_raw"
author: "Jerry Moxley"
date: "2/15/2022"
output: html_document
---
Aim: document database assembly of Caribb reef survey data, primarily raw survey counts of reef fish.  

Database is survey count of species BY country/survey loc/BRUV surveyID.  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
source(here('r', 'fxn_library.R'))

#get the meta data 
uvc <- read_csv(here('data', 'lkup_surveyMethods_uvc.csv'))
```

## Belize

Data cleaning is done for Glovers Reef and SWCMR. 
Needs file wrangling into overall pipeline. 
lkup_Blz_areaStandardization_Jan22 is db structure src.  

```{r cars}
summary(cars)
```

## Colombia
30x2m Tubes
Only a subset of families monitored (see screenshot)

Islas Rosario 
FP 478_2016 at Isla Grande (10.1784107	-75.75384933)
FP 479_2016 at Isla Tesoro (10.23514	-75.74047)
6 other "b" sites surveyed in Rosario archipel

Islas Bernardo
FP480_2016 at Tintipan (2 uvc depths avail).. (9.7858667	-75.7885083)
FP481_2016 at Mangle (2 uvc locs avail)..     (9.75363335	-75.796425)
6 other "b" sites surveyed in Bernardo archipel


```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
