---
title: "Traits EDA"
author: "Jerry Moxley"
date: "5/30/2022"
output: html_document
---
Doc to track analytical decision in a traits-based analysis of reef fish data in the Caribbean.  

May 2022, currently working with subset of reef fish data b/c of issues around integrating cylinder-surveys/radial transects of PR/FlKeys data.  

```{r setup, include=FALSE}
##Workbench set up
library(tidyverse)
library(here)
source(here('r', 'stitch_db_pipe.R')) #data & traits db
source(here('r', 'fxn_library.R'))
```

## Density calcs
effort units are encoded in the 'eff.' fields.

data details for calculating density at site_reef scales
Belize - src reports # of a spp within a size class viewed on a transect (most sites surveyed 8 times).. 

density calc'ed = sum(spp count for site_reefcode) / (eff_nsrvyed * eff_pue) for each of the 2 site-zones (Glovers, SWC)

Cuba - src reports # of spp observed on all 10 transects at a site_reef 

density calc'ed = count / (eff_nsrveyed * eff_pue)

Colombia - src reports  # of spp sighted on reef transect w/in site_reefs w/in site zones (rosario, bernado)

density calc'ed = sum(spp count for each site_reef (or 1st 3 letters of reef_code)) / eff_nsrveyed * eff pue)

Caicos - src reports # of spp sighted for a survey date at sight.  transect field is a numeric encoding of date.  site reef code is site // season // year
also has staff/student observers

density calc'ed = sum(spp count for each site_reef_code)



```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
