---
title: "Wrgl_FishCounts_raw"
author: "Jerry Moxley"
date: "2/15/2022"
output: html_document
---
Aim: document database assembly of Caribb reef survey data, primarily raw survey counts of reef fish.  

Database is survey count of species BY country/survey loc/BRUV surveyID.  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
source(here('r', 'fxn_library.R'))
```

##effort DB Initialize
Initalize an database for tracking survey effort across methods
```{r}
#initalize lkup_surveyeffort from BRUV data
  #initalize lkup_surveyeffort from BRUV data
srvy <- read_csv(here('data', 'summStats_tbl_4.csv')) %>% 
    janitor::clean_names() %>%
    select(reef_id, trip_year, location_name, site_name, reef_name) %>% 
    mutate(
           ##sampling site meta data
           country = location_name,
           site.zone = site_name, 
           site.reef = paste(site_name, reef_name, trip_year, sep = "_"),
           site.reefcode = reef_id, 
           lat = NA, lon = NA, year = trip_year,
           
           ##srvy Params
           srvy.type = "bruv", #OR uvc.f[ish] uvc.b[enthic]
           srvy.method = "cam", #OR radial[uvc], belt[uvc], quad[rat]
           srvy.taxa = "elasmo", #elasmo, fish, benthos
           d2bruv = NA, #dist to bruv [unit]
           fpid = NA, #assigned bruv id based on proximity
           n.obs = NA, #n of samples in src data
           
           ##effort params
           eff.nsamples = NA, #number of transect samples
           eff.nsrvyed = NA, #number of srveyed 
           eff.pue = NA, #number for PUE standardization
           eff.unit = NA, #m2 or perMin(??)
           
           ##file tracking mgmt
           #is it stitched in[t/f], raw src [file], raw out [file]
           stitch.ed = NA,
           stitch.in =  "BRUVmaxn_elasmobranch_observations_all.csv", 
           stitch.out = "src_BRUV_westernAtl.csv",
           dat.partner = "finprint"
           
           ) %>% 
    select(-location_name, -site_name, -reef_id, -trip_year, -reef_name)
#write-out srvy data form
write_csv(srvy, here('data', 'lkup_srvy_params.csv'))

###BRUV sweep
bruv <- read_csv(here('data/stitch', 'src_BRUV_westernAtl.csv'))
  # merge(
  #   #merge bruv data here
  #   #or do in separate wrgnl process.. <<<- probs this!
  # )
  #   write_csv(here('data', 'lkup_surveyeffort.csv')
  # )


####dB Initialize.. v.1
dB <- tibble(country = NULL,
             fish = NULL,
             uvc.f = NULL)
#####initialize dB w/ master.data.frames
uvc.f.cols <- c(
  "site.reef", "site.zone", "site.reefcode",
  "year", "country", "sci.name", "count", 
  "transect","eff.nsites", "eff.nsrvyed", "eff.pue")
dF.f <- data.frame(NULL)
```
Follow up
wrangle bruv effort into lkup table
change eff.nsites to eff.nsamples or eff.ntransects

## Belize

Data cleaning is done for Glovers Reef and SWCMR. 
Needs file wrangling into overall pipeline. 
lkup_Blz_areaStandardization_Jan22 is db structure src.  

RAW DATA has size classes, if size mediated bc impt.
SRC DATA is from initial dens ests; hand copies of sheets in wintr 22.


EFFORT
lkup_belize_areaStandardization.csv shows sites where data was only inputted for fewer than total of that's reef's yrly 8 transects.  

Default effort is 8 transects/site w/ 60m tubes.. coded by max(transect) for a site/year pair.     

No lat/lons currently.

site.zone defined as Glovers/GRMR or SWCMR.  Site.reef and Site.reef code are the same, but maybe should use designations baked into site code.  

```{r cars}
#wrangle
source(here('r', 'wrngl_uvc_belize.R'))

#LOTTA LOTTA FOLLOW UPS HERE.  

#store
#listcol approach
dB <- dB %>% bind_rows(
    tibble(country = "belize", 
      fish = list(out$fish.belize),
      uvc.f  = list(out$uvc.f.effort.belize)
    ))
#df-nesting approach
dF.f <- dF.f %>% 
  bind_rows(out$fish.belize %>% 
              select(uvc.f.cols))
```
follow up.. 
- lil confusd on assignment of site reef & site reef code.  I believe src$site is for an individual reef.  Not sure how to assign zone.  On 2/17, site and site.reef & site.reef.code are assigned the same id src>site.  Eval if should be the locale baked into the reef cof
- consider pre summary of belize data (??)
- needs spp codes incorporated
- for loop needs water tightening 
- no lat/lons
- Should the agg over size classes be done prior to stitch/src data?  
- do you want remaining fields from lkup_samplingSites
- DOUBLE CHECK MERGE ON SITES & RESERVES.. THERE ARE MIRRORING SITE CODES BW RESERVE ZONES.  I **THINK** they are in opposing odd/even years tho.  - dupes in the effort database??
- get spp.ncodr working
- DOUBLE DOUBLE CHECK ASSINGMNET OF MERGERS wrt to year/site/zone pairs
- I thnk technically I should move the new field definition out of the for loop into the df call.. double chk
-DUPES IN THE DF.. def a problem with reserve assignment.  Re-eval!!

## Colombia
30x2m Tubes (in simac metodos.pdf)
Somera and Media depths (ignoring this distinction rn)
Only a subset of families monitored (see screenshot)

Islas Rosario 
FP 478_2016 at Isla Grande (10.1784107	-75.75384933)
FP 479_2016 at Isla Tesoro (10.23514	-75.74047)
6 other "b" sites surveyed in Rosario archipel

Islas Bernardo
FP480_2016 at Tintipan (2 uvc depths avail).. (9.7858667	-75.7885083)
FP481_2016 at Mangle (2 uvc locs avail)..     (9.75363335	-75.796425)
6 other "b" sites surveyed in Bernardo archipel

code for avg locs
bruv %>% filter(location_name == "Colombia") %>% filter(reef_id == "481") %>% summarize(lat = mean(set_lat), lon = mean(set_long)) %>% unlist()
```{r pressure, echo=FALSE}
source(here('r', 'wrngl_uvc_colombia.R'))
#calc density at basal level

#arrange effort according to srvy fields
#effort %>% select(colnames(srvy)) %>% colnames

##bind dat w/ relevant master df's
dB <- dB %>% bind_rows(
    tibble(country = "colombia", 
      fish = list(out$fish.colo),
      uvc.f  = list(out$uvc.f.effort.colo)
    ))
#df-nesting approach
dF.f <- dF.f %>% bind_rows(out$fish.colo %>% select(uvc.f.cols))
```

Follow ups:
-Numbers in ptnr-provided "Transectos por estaciones de monitoreo" do not match data src. 
-still need to bind outputs into the master df
- get spp encoder going into 'spp' field

Wrangle debrief: 
Straight-forwarded, raw provided in spp counts X transect/site/region.  One encoding error with a spp.  60m2 belt transects  only recorded subset of "selected spp" (may be quasi-citSci data).  Can be sampled at a per-transect basis, as is.  

##  Cuba
Values provided are densities, back-calculated to counts via effort info seen below (and in .xlsx of raw data).
Biomass data also available, ignored for now. 

From raw data
16 sites on reef slope 15-20 m deep in 2018.
10 belt transcts 50 m x 4 m per site.
Species list includes large fishes (target by fisheries), herbivorous and medium size fishes, mainly urchin predators.
Data are density per site on individuals/100 m2

```{r}
source(here('r', 'wrngl_uvc_cuba.R'))

#wrangle effort into srvy table
dB <- dB %>% bind_rows(
    tibble(country = "cuba", 
      fish = list(out$fish.cuba),
      uvc.f  = list(out$uvc.f.effort.cuba)
    ))
#df-nesting approach
dF.f <- dF.f %>% 
  mutate(transect = as.character(transect)) %>% 
  bind_rows(out$fish.cuba %>% 
              select(uvc.f.cols))
```
Follow up:
- need to develop wrangling of the effort for lkup table.  
- make site ref codes
- could build easy sampler for bootstrapping transects

##Caicos
- reported in density data, back calc' to spp count. 
- uvc tube is 2x100 acc to Ruth's methods (actually seems tube dims are reported alongside the data).. (ACCORDING TO EMAIL, IT'S A 4X2X100M TUBE)
- has designated student VS staff-conducted surveys, structured seasonally/yrly

-only input surveys w/ counts reported (i.e., filter out sampling like (single, few, many, abundant)
-has trophic assignment based on chaplin paper, may be worthwhile comparison w/ paddack assignments
- does have size data
- don't know what the time field is about
- has depth data
- effort sites are based on site name & season (date surveyed is unique to season, acc. to this code: dat %>% group_by(year, season, site) %>% distinct(date)


- vast majority of 2016 surveys are student-leds
- student obs appear limited to 


```{r}
#source script here
source(here('r', 'wrngl_uvc_caicos.R'))


dB <- dB %>% bind_rows(
    tibble(country = "caicos", 
      fish = list(out$fish.caicos),
      uvc.f  = list(out$uvc.f.effort.caicos)
    ))
#df-nesting approach
dF.f <- dF.f %>% 
  bind_rows(out$fish.caicos %>% 
              mutate(year = as.numeric(year),
                     transect = as.character(transect)) %>% 
              select(uvc.f.cols))

```

- needs follow up on site assignment & which sites are valuable samplers for South Caicos reserve
- site.zone is hard-coded as "South Caicos"
- has some rando site names (apply to surveys in 2014-2015)
- might require someway to reconcile date of survey & a transect index.. impacts how nsites & nsrveys is calc'ed (which is hella important)
- i don't understand "transect" as a field

-FAIRLY GOOD EMAIL CORRESPONDENCE THAT SHOULD BE MINED AND ALIGNED
- decision necessary about # of years to include, w/ finprint data
- decision about how to handle season data? (as Feb 22, site.reefcode has season baked into it)
-transect is set as date, could be indexed

-maybe quick analysis comapring student- v staff-led?
-Apr22, filtering for staff only 

-EFFORT df needs big attn!! (both in the summarization, and in the merger w/ df)
- Evaluate the duplicates in the spp data

##FLORIDA
radial transects.. lots of zeros, extensive spp list. 
src field desc: https://grunt.sefsc.noaa.gov/rvc_analysis20/samples/index

src data is entirely 2016, but applies to BRUVS sampling in 2016 (4 surveys) and 2017 (n = 2 srvys) BRUV sr

could pull 2017 data via curl script (in data/raw)?

potential for taxa list as a lkup tbl?

src data has length & num fields.. but num are doubles, look kinda like densities (some massive numbers too??)
acc to data diction:
LEN: The length, in cm, of a sample.
NUM: The number of individuals of a given species and length observed in a secondary sampling unit. (nB: REPORTED AS AN AVG COUNT W/IN DIVE BUDDY PAIR)
STATION_NR: code assigning field to survey by first or second buddy team

effort structure
ctry - 'florida'
zone - subregion_nr [8:12] (applies to uppper/lower keys?)
reef - primary sampling unit
reefcode - PSU + date surveyed
transect - (secondary) station number (ie. 2 radial transects per sample)

habitat codes defined as: 
(from slide6: https://nmsfloridakeys.blob.core.windows.net/floridakeys-prod/media/archive/sac/othermaterials/20150616rvc.pdf) <-- pretty helpful ppt on sampling/estimation
-offshore patch reef
-mid-channel patch reef
-back reef/rubble
-low relief spur/groove
-high releif spur/groove
-patchy harbottom
-low relief har bottom
-

keys/pr RVC methodology published here: https://www.sciencedirect.com/science/article/pii/S0165783611000312
```{r}
#source script here
source(here('r', 'wrngl_uvc_florida.R'))

dB <- dB %>% bind_rows(
    tibble(country = "florida", 
      fish = list(out$fish.fla),
      uvc.f  = list(out$uvc.f.effort.fla)
    ))
#df-nesting approach
dF.f <- dF.f %>% 
  bind_rows(out$fish.fla %>% 
              mutate(transect = "NEEDSattn") %>% 
              select(uvc.f.cols))

```
methods details for data collection: https://nsuworks.nova.edu/cgi/viewcontent.cgi?article=1093&context=occ_facreports/
taxa uses NOAA-provided taxa codes (eval using this tbl for rest of dataset)

-ATTN TO EFFORT METRICS & ENSURE SAMPLING HIERARCHY IS ACCRUATELY AGGREGATED
- might need add'l group-by statements depending on aggregaton scale ^^
******
- EFFORT ZONED BY SUBREGION [8:12] & hab.code (APR22, double check w/ jBlondeau)
- ZONES AGGREGATED BY site.reef (reef idx) AND site.reefcoe (reefIdx + date surveyed)
******
-what to do w/ habitat codes, maritime zone (inshore, offshore, midchannel, forereef)
- 2017 data pull???

this overall fish density provides a DQ-chk for density ests: 
Total mean density for all sites and strata combined for all three years was 170 ±5.9 SEM
fishes/SSU. (https://nsuworks.nova.edu/cgi/viewcontent.cgi?article=1093&context=occ_facreports/)

##Puerto Rico

BRUV sampling is 2017; raw PR folder catalogues uvc from 2016 & 2019. 

Largely, same data fields/sampling hierarchy as Florida.. and shld follow near same pre-processing. 
e.g., uses FLA taxa list for species codes

IMPT distinction: 
PR only uses 1 dive buddy team (ie., only 1 station_nr per PSU)
FLA uses 2 dive-buddy teams for counts (NUM is the avg of a buddy team's individual counts; station_nr {i presume} are the counts reported for each team.  )
see here: https://www.nodc.noaa.gov/archive/arc0101/0157633/8.8/data/0-data/Atlantic/Biological/Florida/NCRMP_RVC_Protocol_2017.pdf
```{r}
###CURRENTLY TAKING AN UNWIELDY AMT OF TIME
source(here('r', 'wrngl_uvc_pr.R'))

# dB <- dB %>% bind_rows(
#     tibble(country = "pr", 
#       fish = as_tibble(out$fish.pr),
#       uvc.f  = as_tibble(out$uvc.f.effort.pr)
#     ))
#df-nesting approach
dF.f <- dF.f %>% 
  bind_rows(out$fish.pr %>% 
              select(uvc.f.cols))
```

-EFFORT zoned by habitat codes, aggregated by site reefs
-need JBlondeau insight on sample aggregation

-determine application of 2019 data
-confirm same sampling area in cylinder
-needs some more definition of effort statistics

##GUADELOUPE

UVC fish is absent for Guadeloupe. 
Follow up RDunn re: absent data?

6 Benthic transects avail in folder. 
Use RDunn estimates.  

##DB STITCH QC
-lots of zeros
```{r}
sum(is.na(dF.f$count)) #8 NA's
dF.f <- dF.f %>% filter(!is.na(count))

sum(dF.f$count == 0, na.rm = T) #2/3rds of data are zeros
dF.f %>% 
  filter(count == 0) %>% 
  group_by(country) %>% 
  count()
#98% of them are NOAA

#going to remove zeros
dF.f <- dF.f %>% 
  filter(count != 0)

dF.f %>% summary()
```


##TROPHIC ASSIGNMENTS
Species are assigned trophic groupings & functional roles based on lookup tables. 

Sources for lkups are:
-Paddack et al. 
-NOAA florida table
-diaz fxnl traits
-micheli ?? 
-paddack ref w/ hab-generalist/specialist (4 spp w/o representation in og paddack tbl)

-anything to draw from here: https://www.pnas.org/doi/10.1073/pnas.2012318118
-hab specialist/generalist designated here (see Luiz ref): https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0126004#pone.0126004.s008
-also provides rates of change in spp abundance trends (gen. rel. little change)

diaz groupings (multiple assignment, all fields)
- water col feeding position (wcfp): benthic (BEN), demersal (DEM), midwater (MID), and surface (SUR)
- social feeding behavior (sfb): soliatry indis (SOL), in pairs (PAIR), intraspecific aggregations (AGR), or in mixed groups with other species (MIX)
- Feeding time (ftod): diurnal (DIU), crepuscular (CRE), and nocturnal (NOC), 
- feeding habitat: coral reefs (COR), rocky reefs (ROC), seagrasses(SEA), algae beds (ALG), mangroves (MAN), sand (SAN), coral rubble (RUB), open water (OPE), water column (COL), out of shelf (SHE), and attached to other verts (VER)

2 continuous/ordinal  variables
-TL - 2.0 (herbs) up to 4.5 (piscivores)
- diet breadth - numbe of prey idtems included in deit (nekton, zooplankton, zoobenthos, detritus, plants, sand, organic matter, ectopnarasites, and unidentified animal material)


habitat specialization
4 spp not in src paddack tbl: 
[1] "Elacatinus illecebrosus"   
[2] "Haemulon vittatum"         
[3] "Kyphosus incisor/sectatrix"
[4] "Calamus pennatula"  

```{r}
#PADDACK MERGE
 sc <- dF.f %>% 
  mutate(sci.name = gsub("_", " ", sci.name)) %>% 
  mutate(sci.name = tolower(sci.name)) %>% 
  merge(read_csv(here('data','lkup_trophic_akk.csv')) %>% 
  janitor::clean_names() %>% 
  mutate(sci.name = 
           tolower(paste(genus, species, sep = " "))) %>% 
  select(sci_name, family, max_length, 
         tg_pdk = trophic_group, 
         fished_pdk = fishing_status, 
         habspec_pdk = habitat_use), 
  by = "sci.name", all.x = T) %>% 
  mutate(
    tg_pdk = as.factor(trophic_group),
    fished_pdk = as.factor(fished_pdk),
    habspec_pdk = as.factor(fished_pdk)
    family = as.factor(family)
  ) %>%
  #filter some spurious vals
  filter(sci.name != "#n/a") %>% 
  filter(!is.na(sci.name))

#LOOK INTO NAs
dF.f %>% group_by(country) %>% summarize(nNA = sum(is.na(trophic_group)))
#LOT OF FLORIDA NA'S

##EDA PLOT
dF.f %>% 
  group_by(country, site.reefcode, trophic_group) %>% 
  summarize(
    dens = sum(count, na.rm =T)/eff.pue
      )

##NOAA MERGE (spp codes, )
```

 Follow up
 - check NA matches of sci.names.. add to Paddack?

##FULL FOLLOW UP
 - determine what to do with the integration fields
 - get field names straight bw . & _.. particularly if janitor::clean_names((0)) is gettitng called. 
 - is n.obs a useful metric at all? probably worthwhile compiling this at the end when data is agg'ed
 - wont data df's need a zone col?
 - get spp.ncodr working.
 
 ##OTHER DATA INTG
 
 - MVermj contributed reef fish densities to Sandin et al. 2022 Emerg Top Life Sci 6 (1): 57–65.  https://portlandpress.com/emergtoplifesci/article/6/1/57/230986
 -No BRUV samples for Curacao/Bonaire
 - FxnDiversity paper might contribute data from USVI (uses belt transects..).. see program here: https://coastalscience.noaa.gov/project/caribbean-coral-reef-ecosystem-monitoring-project/
